---
title: "Analysis Pipeline"
author: "Nicholas Borcherding"
date: '2022-07-11'
output: html_document
---

###################
#Processing Data
#################

General pipeline for the generating intermediate Seurat objects that have been 1) filtered/passe qc and 2) anntoated with multiple packages

```{r eval=FALSE}
dir.create("dataExplore")
dir.create("dataExplore/qc")
"%!in%" <- Negate("%in%")
file_list <- list.files("./data/")

library(SingleR)
library(Matrix)
library(ProjecTILs)

#############################
#Laoding Annotation Data Sets
############################

DICE <- DatabaseImmuneCellExpressionData()
ref8 <- load.reference.map("./data/annotations/ref_LCMV_Atlas_mouse_v1.rds")
ref4 <- load.reference.map("./data/annotations/ref_LCMV_CD4_mouse_release_v1.rds")
    
for (i in seq_along(file_list)){
    tmp <-  Read10X(paste0("./data/sequencingRuns/", file_list[i], "/filtered_feature_bc_matrix"))
    s.obj <- CreateSeuratObject(tmp, project = file_list[i])
    rm(tmp)
    s.obj <- RenameCells(object = s.obj, new.names = paste0(file_list[i], "_", rownames(s.obj[[]])))
    s.obj<- subset(s.obj, subset = nFeature_RNA > 100)
    s.obj[["mito.genes"]] <- PercentageFeatureSet(s.obj, pattern = "^MT-")
    
    p1 <- VlnPlot(object = s.obj, features = c("nCount_RNA")) + theme(legend.position = "none")
    p2 <- VlnPlot(object = s.obj, features = c("nFeature_RNA")) + theme(legend.position = "none")
    p3 <- VlnPlot(object = s.obj, features = c("mito.genes")) + theme(legend.position = "none")
    
    pdf(paste0("./dataExplore/qc/", file_list[i], ".pdf"), height = 8, width=12)
    grid.arrange(p1, p2, p3, ncol = 3)
    dev.off()
    
  ###########################
  #Here is the filtering step
  ############################
  standev <- sd(log(s.obj$nFeature_RNA))*2 #cutting off above standard deviation of 2
  mean <- mean(log(s.obj$nFeature_RNA))
  cut <- round(exp(standev+mean))
  
  p4 <- FeatureScatter(s.obj, feature1 = "nCount_RNA", feature2 = "mito.genes") + 
      geom_hline(yintercept = 15)
  p5 <- FeatureScatter(s.obj, feature1 = "nCount_RNA", feature2 = "nFeature_RNA") + 
      geom_hline(yintercept = cut)
   
  pdf(paste0("./dataExplore/qc/", file_list[i], "_scatterPlot.pdf"), height = 4, width=8)
    grid.arrange(p4, p5, ncol = 2)
  dev.off()
  
  s.obj <- subset(s.obj, subset = mito.genes < 15 & nFeature_RNA < cut)
    
  ###########################################
  #Estimate Doublets for Each Sequencing Run
  ############################################
  sce <- as.SingleCellExperiment(s.obj)
  sce <- scDblFinder(sce, BPPARAM=MulticoreParam(3))
  doublets <- data.frame(db.weight.score = sce$scDblFinder.score, db.ratio = sce$scDblFinder.weighted, 
                         db.class = sce$scDblFinder.class, db.score = sce$scDblFinder.score)
  rownames(doublets) <- rownames(sce@colData)
  s.obj <- AddMetaData(s.obj, doublets)
  rm(sce)
  
  ## Finding variable features and normalizing data for integration to follow
  s.obj <- NormalizeData(s.obj)
  s.obj <- FindVariableFeatures(s.obj, selection.method = "vst", nfeatures = 2000)
  s.obj <- quietTCRgenes(s.obj)
  
  #############################################
  #Projectil Annotation of Cell Types
  #############################################
  
    query.projected <- make.projection(s.obj, ref = ref4, ncores = 1)
    
    #Lapply across list of Seurat objects
    query.projected <- cellstate.predict(ref = ref4, query =query.projected , reduction = "umap", ndim = 2)
      meta <- query.projected[[]][c("functional.cluster", "functional.cluster.conf")]
      colnames(meta) <- c("CD4.annot", "CD4.score")
      rownames(meta) <- stringr::str_remove(rownames(meta), "Q_")
    s.obj <- AddMetaData(s.obj, meta)
    rm(query.projected)

    ###########################
    #CD8 LCMV Annotation
    ############################
    
    query.projected <- make.projection(s.obj, ref = ref8, ncores = 1)
    
    query.projected <- cellstate.predict(ref = ref8, query = query.projected, reduction = "umap", ndim = 2)
    meta <- query.projected[[]][c("functional.cluster", "functional.cluster.conf")]
    colnames(meta) <- c("CD8.annot", "CD8.score")
    rownames(meta) <- stringr::str_remove(rownames(meta), "Q_")
    s.obj <- AddMetaData(s.obj, meta)
    rm(query.projected)
    rm(ref)
    
    
    DICE.annotation <- NULL
    tmp.2 <- s.obj@assays[["RNA"]]@counts
    ####This approach for matrix conversion saves some memory
    tmp.2 <- tmp.2[tabulate(summary(tmp.2)$i) != 0, , drop = FALSE]
    com.res2 <- SingleR(tmp.2, ref=DICE, labels=DICE$label.fine, assay.type.test=1)
      df <- data.frame("DICE.labels" = com.res2$labels, "DICE.pruned.labels" = com.res2$pruned.labels)
      rownames(df) <- rownames(s.obj[[]])
      s.obj <- AddMetaData(s.obj, df)
 saveRDS(s.obj, file = paste0("./data/", file_list[i], "/processedRNA_SeuratObject.rds"))
 rm(s.obj)
}
```

###############
#Integrate Data
###############

```{r}
library(Seurat)
library(dplyr)
library(harmony)
library(scRepertoire)

inventory <- read.delim("./data/data.inventory.txt")
samples<- inventory$donor

list <- list()
for (i in seq_along(samples)) {
  list[[i]] <- readRDS(paste0("./data/sequencingRuns/", samples[i], "/processedRNA_SeuratObject.rds"))
}

#Merging Data
list <- merge(list[[1]], list[-1])

data.inventory <- read.delim("./data/data.inventory.txt")
meta <- list[[]]
#meta <- meta[,-which(colnames(meta) %in% colnames(data.inventory))]
barcodes <- rownames(meta)
meta <- merge(meta, data.inventory, by.x = "orig.ident", by.y = "sample")
meta <- meta[,colnames(data.inventory)[-2]]
rownames(meta) <- barcodes

list <- AddMetaData(list, meta)


list <- list %>%
    Seurat::NormalizeData(verbose = FALSE) %>%
    FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>% 
    ScaleData(verbose = FALSE) %>% 
    RunPCA(pc.genes = pbmc@var.genes, npcs = 30, verbose = FALSE) %>%
    RunHarmony(c("orig.ident")) %>%
    FindNeighbors(reduction = "harmony", dims = 1:25) %>%
    RunUMAP(reduction = "harmony",
            dims = 1:25,
            n.epochs = 500) %>%
    FindClusters(resolution = 1, 
                 algorithm = 3) %>% 
    identity()

saveRDS(list, file = "./data/processedData/Fullintegrated_seuratObjects.rds")
```

###################
#Adding Contig Data
###################

```{r}
library(scRepertoire)

######################################
#iterate to make a list of contig csvs
######################################


contig.list <- NULL
for (i in seq_along(samples)){
  contig.list[[i]] <-  read.csv(paste0("./data/sequencingRuns", samples[i], "/filtered_contig_annotations.csv"))
}
names(contig.list) <- samples

##################################################
#Reducing the data to the individual barcode level
##################################################
combinedObject <- combineTCR(contig.list, samples = samples, filterMulti = TRUE, removeNA = TRUE, cells = "T-AB")

pt <- data.inventory$donor[match(samples, data.inventory$sample)]
combinedObject <- addVariable(combinedObject, name = "Patient", variables = pt)
list <- combineExpression(combinedObject, list, cloneCall = "aa", group.by = "Patient")

slot(list, "meta.data")$cloneType <- factor(slot(list, "meta.data")$cloneType, 
                levels = c("Hyperexpanded (0.1 < X <= 1)", 
                           "Large (0.01 < X <= 0.1)", 
                            "Medium (0.001 < X <= 0.01)", 
                            "Small (1e-04 < X <= 0.001)", 
                            "Rare (0 < X <= 1e-04)", NA))

DimPlot(list, group.by = "cloneType") + scale_color_manual(values = rev(viridis_pal(option = "B")(length(unique(list$cloneType)))))

saveRDS(list, file = "./data/processedData/Fullintegrated_seuratObjects.rds")
```


######################
#Iolating Only T Cells
######################

```{r}
#Filtering using the scGate built into PorjecTIL based on T-AB
#Only T cells with contig information
list$filter  <- "Yes"
list$filter <- ifelse(!is.na(list$CD8.annot) | !is.na(list$CD4.annot), "No", list$filter)
list$filter <- ifelse(is.na(list$CTaa), "Yes", list$filter)

list <- subset(list, filter != "Yes")

list@meta.data$CD8.pos <- ifelse(list@assays$RNA@data["CD8A",] >=  0.4, "Yes", "No")

dim(list)[2]
#Check: 219283 cells

list <- list %>%
    Seurat::NormalizeData(verbose = FALSE) %>%
    FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>% 
    ScaleData(verbose = FALSE) %>% 
    RunPCA(pc.genes = pbmc@var.genes, npcs = 30, verbose = FALSE) %>%
    RunHarmony(c("orig.ident", "tissue"), 
               max.iter.harmony = 15) %>%
    FindNeighbors(reduction = "harmony", dims = 1:15) %>%
    RunUMAP(reduction = "harmony",
            dims = 1:15,
            n.epochs = 500) %>%
    FindClusters(resolution = 0.8, 
                 algorithm = 3) %>% 
    identity()

saveRDS(list, file = "./data/processedData/Fullintegrated_seuratObjects_Tcells.rds")


#Relevel the Clonotype Expansion assignments
meta <- list[[]]
meta <- split(meta, meta$donor)

cloneTypes=c(None = 0, Rare = 1e-5, Small = 1e-4, 
                              Medium = 1e-3, Large = 1e-2, Hyperexpanded = 1)

Con.df <- NULL
for(i in seq_along(meta)) {
  tmp <- meta[[i]]
  CDstatus <- table(tmp$CD8.pos)
  tmp <- tmp %>% group_by(tmp[,"CTaa"], CD8.pos) %>%
                      summarise(Clone.Prop = n())
  tmp$Clone.Prop[which(tmp$CD8.pos == "No")] <- tmp$Clone.Prop[tmp$CD8.pos == "No"]/CDstatus[1]
  tmp$Clone.Prop[which(tmp$CD8.pos == "Yes")] <- tmp$Clone.Prop[tmp$CD8.pos == "Yes"]/CDstatus[2]
  colnames(tmp)[1] <- "CTaa"
  data.CD4 <- merge(meta[[i]][meta[[i]]$CD8.pos == "No",], tmp[which(tmp$CD8.pos == "No"),], by = "CTaa")
  data.CD8 <- merge(meta[[i]][meta[[i]]$CD8.pos == "Yes",], tmp[which(tmp$CD8.pos == "Yes"),], by = "CTaa")
  data <- rbind.data.frame(data.CD4, data.CD8)
  data <- data[,c("barcode", "Clone.Prop")]
  Con.df <- rbind.data.frame(Con.df, data)
}

names.cell <- Con.df$barcode
Con.df <- as.data.frame(Con.df[,-1])
rownames(Con.df) <- names.cell
colnames(Con.df) <- "Clone.Prop"
Con.df$Clone.Group <- NA

for (x in seq_along(cloneTypes)) { names(cloneTypes)[x] <- 
        paste0(names(cloneTypes[x]), ' (', cloneTypes[x-1], 
        ' < X <= ', cloneTypes[x], ')') }
    for (i in 2:length(cloneTypes)) { 
      Con.df$Clone.Group <- ifelse(Con.df$Clone.Prop > cloneTypes[i-1] & 
                                Con.df$Clone.Prop <= cloneTypes[i], 
                                names(cloneTypes[i]), Con.df$Clone.Group) 
      }

list <- AddMetaData(list, Con.df)
```

#########################################
#Adding Known Epitope Sequences Annotation
##########################################

```{r}
#Spike 167-180 epitope
sequence <- read.csv("./sequenceInfo/Sequences.csv")
CTaa <- as.data.frame(stringr::str_split(list$CTaa, "_", simplify = TRUE))
TA <-which(CTaa$V1 %in% sequence$cdr3a)
TB <-which(CTaa$V2 %in% sequence$cdr3b)

list$TCRB.S167.180<- NA
list$TCRB.S167.180[TB] <- "Specific"

list$TCRA.S167.180 <- NA
list$TCRA.S167.180[TA] <- "Specific"


#Spike 815-830 epitope

sequence <- read.csv("./sequenceInfo/dykema.sequences.csv")
CTaa <- as.data.frame(stringr::str_split(list$CTaa, "_", simplify = TRUE))
TA <-which(CTaa$V1 %in% sequence$cdr3a)
TB <-which(CTaa$V2 %in% sequence$cdr3b)

list$TCRB.S815.830 <- NA
list$TCRB.S815.830[TB] <- "Specific"

list$TCRA.S815.830 <- NA
list$TCRA.S815.830[TA] <- "Specific"

# VDJdb sequence

sequences <- read.delim("./sequenceInfo/VDJdb.tsv")
sequences <- sequences[sequences$MHC.class == "MHCII" & sequences$Epitope.gene == "Spike",]

epitope <- c("NCTFEYVSQPFLMDL" = "S164.179",
"RISNCVADYSVLYNS" = "S356.371",
"NFSQILPDPSKPSKR" = "S800.815",
"HWFVTQRNFYEPQII" = "S1101.1115",
"VGGNYNYLYRLFRKS" = "S445.459",
"NNSYECDIPIGAGIC" = "S657.671")

list$TCRB.VDJ.db <- NA
list$TCRA.VDJ.db <- NA
for (i in seq_along(epitope)) {
  sequences$Epitope[which(sequences$Epitope == names(epitope)[i])] <- epitope[i]
  TA <- sequences[sequences$Gene == "TRA" & sequences$Epitope == epitope[i], ]
  TB <- sequences[sequences$Gene == "TRB" & sequences$Epitope == epitope[i], ]
  TA.loc <-which(CTaa$V1 %in% TA$CDR3)
  TB.loc <-which(CTaa$V2 %in% TB$CDR3)
  list$TCRB.VDJ.db[TB.loc] <- epitope[i]
  list$TCRA.VDJ.db[TA.loc] <- epitope[i]
}


list$TCRB.epitope  <- NA
list$TCRB.epitope <- ifelse(!is.na(list$TCRB.S167.180), "S167.180", NA)
list$TCRB.epitope <- ifelse(!is.na(list$TCRB.S815.830), "S815.830", list$TCRB.epitope)
list$TCRB.epitope <- ifelse(!is.na(list$TCRB.VDJ.db) & is.na(list$TCRB.epitope), list$TCRB.VDJ.db, list$TCRB.epitope)

list$TCRA.epitope  <- NA
list$TCRA.epitope <- ifelse(!is.na(list$TCRA.S167.180), "S167.180", NA)
list$TCRA.epitope <- ifelse(!is.na(list$TCRA.S815.830), "S815.830", list$TCRA.epitope)
list$TCRA.epitope <- ifelse(!is.na(list$TCRA.VDJ.db) & is.na(list$TCRA.epitope), list$TCRA.VDJ.db, list$TCRA.epitope)

saveRDS(list, file = "./processedData/Fullintegrated_seuratObjects_Tcells.rds")
```


###################
#Figure 1 Graphs
###################

```{r}
dir.create("./output/Figure1/")
list <- readRDS("./data/processedData/Fullintegrated_seuratObjects_Tcells.rds")
```


# Figure 1C
```{r}
DimPlot(list, label = TRUE, raster=FALSE, pt.size = 0.5, repel = TRUE) + 
  theme_void() + 
  scale_color_manual(values = mycolors) 

ggsave("./output/Figure1/Figure1C.pdf")
```

# Figure 1D
```{r}
source("./R/utils.R")
library(Nebulosa)

genes <- c("CD4", "CD8A", "CCR7", "SELL", "CXCR5", "ICOS", "FOXP3", "IL2RA", "CTLA4", "PDCD1")

x <- plot_density(list,
                  features = genes, 
                  size = 0.5,
                  combine = FALSE)

for(i in seq_along(x)) {
  x[[i]] <- x[[i]] + 
              guides(color = "none") + 
              scale_color_gradientn(colors = rev(brewer.pal(11, "RdYlBu"))) +
              theme_void() + 
              theme(plot.title = element_blank(), 
                    plot.margin = unit(c(0.6, 0.15, 0.6, 0.15), "cm"))
}

plot_a_list(x,5,2)
ggsave("./output/Figure1/Figure1D.pdf", height = 5, width = 2)
```

#Figure 1E
```{r}
df <- data.frame(list@reductions$umap@cell.embeddings, list@meta.data)

#Sampling to allow for equal comparisons for density
sampled <- df %>%
  group_by(timepoint) %>%
  sample_n(25000)

sampled$timepoint <- factor(sampled$timepoint, levels = c("d28", "d60", "d110", "d201"))

ggplot() + 
  geom_point(data = sampled, aes(x = UMAP_1, y = UMAP_2), stroke = 0, color = "grey") + 
  stat_density_2d(data = sampled, aes(x = UMAP_1, y = UMAP_2, fill = stat(level)), 
                  geom = "polygon", bins = 25) + 
  facet_grid(.~timepoint) + 
  xlim(-8,12) + 
  ylim(-11,7) + 
  scale_fill_gradientn(colors = rev(brewer.pal(11, "RdYlBu"))) + 
  guides(fill = "none") + 
  theme_void() 

ggsave("./output/Figure1/Figure1E.png", height = 6, width = 28, dpi = 600, bg = "transparent")
```

#Figure 1F
```{r}
sampled <- df %>%
  group_by(tissue) %>%
  sample_n(90000)

sampled$tissue <- factor(sampled$tissue, levels = c("PB", "LN"))
ggplot() +
  geom_point(data = sampled, aes(x = UMAP_1, y = UMAP_2), stroke = 0, color = "grey") + 
  stat_density_2d(data = subset(sampled, tissue == "LN"), aes(x = UMAP_1, y = UMAP_2, fill = stat(level)), 
                  geom = "polygon", bins = 25) + 
  stat_density_2d(data = subset(sampled, tissue == "PB"), aes(x = UMAP_1, y = UMAP_2, fill = stat(level)), 
                  geom = "polygon", bins = 25) + 
  facet_grid(.~tissue) + 
  scale_fill_gradientn(colors = rev(brewer.pal(11, "RdYlBu"))) + 
  xlim(-8,12) + 
  ylim(-11,7) + 
  guides(fill = "none") + 
  theme_void()  + 
  theme(plot.title = element_blank(),
        strip.text.x = element_blank())
ggsave("./output/Figure1/Figure1F.png", height = 6, width = 14, dpi = 600, bg = "transparent")
```

#Figure 1G
```{r}
#Sorting epitopes into either 167-180 or other
other <- c("S356.371" ,"S164.179","S815.830", "S800.815", "S356.371", "S657.671","S1101.1115")

df$epitope <- NA

other <- c("S356.371" ,"S164.179","S815.830", "S800.815", "S356.371", "S657.671","S1101.1115")
df$epitope[c(which(df$TCRB.epitope == "S167.180"), which(df$TCRA.epitope == "S167.180"))] <- "epitope.S167.180"
df$epitope[c(which(df$TCRB.epitope %in% other), which(df$TCRA.epitope %in% other))] <- "other.spike"

ggplot() +
  geom_density2d(data = df, aes(x = UMAP_1, y = UMAP_2), color = "black", lwd = 0.5) + 
  geom_point(data = subset(df, !is.na(epitope)), aes(x = UMAP_1, y = UMAP_2, fill = epitope), shape=21, stroke = 0.25, size = 3) + 
  guides(fill = "none") + 
  theme_void() + 
  scale_fill_manual(values = rev(brewer.pal(11, "RdYlBu")[c(2,10)]))
ggsave("./output/Figure1/Figure1G.png", height = 6, width = 7, dpi = 600, bg = "transparent")
```

#Figure 1H
```{r}
library(ggseqlogo)
library(msa)

meta <- list[[]]
meta <- meta[!is.na(meta$TCRB.S167.180) |!is.na(meta$TCRA.S167.180),] 
TRB <- unique(str_split(meta$CTaa, "_", simplify = TRUE)[,2])
TRA <- unique(str_split(meta$CTaa, "_", simplify = TRUE)[,1])

#################
#S167-180 Epitope
#################

align <- msa(TRA, 
              type = "protein", 
              maxiters = 30)
  align2 <- msaConvert(align, type="seqinr::alignment")
  
  ggseqlogo(align2$seq, 
            seq_type='aa', 
            method='p') 
  ggsave("./output/Figure1/Figure1H.Specific167.TRA.pdf", height =3, width = 4.5)
  
    align <- msa(TRB, 
               type = "protein", 
               maxiters = 30)
  align2 <- msaConvert(align, type="seqinr::alignment")
  
  ggseqlogo(align2$seq, 
            seq_type='aa', 
            method='p') 
  ggsave("./output/Figure1/Figure1H.Specific167.TRB.pdf", height =3, width = 4.5)

#####################
#Other Spike Epitopes
#####################
  
meta <- list[[]]
meta <- meta[!is.na(meta$TCRA.epitope) & is.na(meta$TCRA.S167.180) |!is.na(meta$TCRB.epitope) & is.na(meta$TCRB.S167.180),] 
  TRB <- unique(str_split(meta$CTaa, "_", simplify = TRUE)[,2])
  TRA <- unique(str_split(meta$CTaa, "_", simplify = TRUE)[,1])
  
  align <- msa(TRA, 
               type = "protein", 
               maxiters = 30)
  align2 <- msaConvert(align, type="seqinr::alignment")
  
  ggseqlogo(align2$seq, 
            seq_type='aa', 
            method='p') 
  ggsave("./output/Figure1/Figure1H.Other.TRA.pdf", height =3, width = 4.5)
  
    align <- msa(TRB, 
               type = "protein", 
               maxiters = 30)
  align2 <- msaConvert(align, type="seqinr::alignment")
  
  ggseqlogo(align2$seq, 
            seq_type='aa', 
            method='p') 
  ggsave(paste0("./output/Figure1/Figure1H.Other.TRB.pdf"), height =3, width = 4.5)
```

##################
#Figure 2 Graphs
##################

```{r}
dir.create("./output/Figure2/")
```

```{r}
list <- readRDS("./data/processedData/Fullintegrated_seuratObjects_Tcells.rds")

tfh <- subset(list, tissue == "LN" & CD4.annot %in% c("Tfh_Memory", "Tfh_Effector"))
```



```{r}
DefaultAssay(tfh) <- "RNA"

tfh <- tfh %>%
    Seurat::NormalizeData(verbose = FALSE) %>%
    FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>% 
    quietTCRgenes() %>%
    ScaleData(verbose = FALSE) %>% 
    RunPCA(pc.genes = pbmc@var.genes, npcs = 30, verbose = FALSE) %>%
    RunHarmony(c("orig.ident")) %>%
    FindNeighbors(reduction = "harmony", dims = 1:25) %>%
    RunUMAP(reduction = "harmony",
            dims = 1:20,
            n.epochs = 500) %>%
    FindClusters(resolution = 0.5, 
                 algorithm = 3) %>% 
    identity()

saveRDS(tfh, "./data/processedData/Fullintegrated_seuratObjects_TFH_TFHonly.rds")
```

#Figure 2A
```{r}
tfh <- readRDS("./data/processedData/Fullintegrated_seuratObjects_TFH_TFHonly.rds")
library(RColorBrewer)
library(ggplot2)
library(Seurat)
library(dplyr)
mycolors <- colorRampPalette(brewer.pal(11, "Paired"))(length(unique(tfh$seurat_clusters)))

update_geom_defaults("point", list(stroke = 0.25))
DimPlot(tfh, reduction = "umap", label = TRUE, repel = TRUE, label.size = 4, raster=FALSE, pt.size = 0.25) +
  scale_color_manual(values = mycolors) + 
  guides(colors = "none", fill = "none") + 
  theme_void() + 
  NoLegend() 

ggsave("./output/Figure2/Figure2A_TFHonly.pdf", height = 3.5, width = 4, dpi = 600, bg = "transparent")
```

```{r}
library(patchwork)
plot <- NULL
features <- sort(c("BCL6", "CXCR5", "CD274", "ICOS", "CCR7","CD40LG","IL21","CXCL13", "IL4", "FOXP3", "IL10", "SELL"))
for (i in seq_along(features)) {

  if(i == 1) {
    plot <- plot_density(tfh,
                features = features[i], 
                size = 0.5) + 
     scale_color_gradientn(colors = rev(brewer.pal(11, "RdYlBu"))) + 
   guides(fill = "none", color = "none") + 
      theme_void() + 
      theme(plot.title = element_blank(), 
            plot.margin = margin(10,1,1,1))
  } else {
    plot <- plot + plot_density(tfh, 
                features = features[i], 
                size = 0.5) + 
     scale_color_gradientn(colors = rev(brewer.pal(11, "RdYlBu"))) + 
    guides(fill = "none", color = "none") + 
      theme_void() + 
      theme(plot.title = element_blank(), 
            plot.margin = margin(10,1,1,1))
  }
}

plot + plot_layout(nrow = 3)
ggsave("./output/Figure2/Figure2B_TFHonly.png", height = 5.3, width = 8, dpi = 600, bg = "transparent")
```

#Figure 2C
```{r}
markers <- FindAllMarkers(tfh)
write.csv(markers, "./output/tfh_onlymarkers.csv")
```


```{r}
markers <- read.csv("./output/tfh_onlymarkers.csv")

top <- markers %>%
  group_by(cluster) %>%
  slice_max(order_by = avg_log2FC, n = 8)

DotPlot(tfh, features = unique(top$gene)) + 
  scale_color_gradientn(colors = rev(brewer.pal(11, "RdYlBu"))) + 
  scale_y_discrete(limits=rev)+ 
  coord_flip()
ggsave("./output/Figure2/Figure2C_TFHonly.pdf", height = 20, width = 6)
```

#Figure 2D
```{r}
ES <- readRDS("./data/ProcessedData/enrichmentResults.rds")

meta <- tfh[[]]
meta$enrichment.groupings <- NA
meta$enrichment.groupings[which(meta$seurat_clusters %in% c(0,1,2,6,7,10))] <- "Memory.TFH"
meta$enrichment.groupings[which(meta$seurat_clusters %in% c(3))] <- "Effector.TFH"
meta$enrichment.groupings[which(meta$seurat_clusters %in% c(4))] <- "TFR"
meta$enrichment.groupings[which(meta$seurat_clusters %in% c(8))] <- "IL10.TFH"
meta$enrichment.groupings[which(meta$seurat_clusters %in% c(11))] <- "Proliferating.TFH"
meta$enrichment.groupings[which(meta$seurat_clusters %in% c(9))] <- "Cytolytic.TFH"
meta <- meta[which(!is.na(meta[,"enrichment.groupings"])),]

ES2 <- ES[rownames(ES) %in% rownames(meta),]
variances <- apply(ES2, 2, var)
#ES2 <- ES2[,-which(variances == 0)]

pathways <- colnames(ES2)

ES2 <- merge(ES2, meta, by = 0)

sig.results <- getSignificance(ES2, 
                               group = "seurat_clusters", 
                               gene.sets = pathways, 
                               fit = "KW")

sig.results2 <- getSignificance(ES2, 
                               group = "seurat_clusters", 
                               gene.sets = pathways, 
                               fit = "ANOVA")

KW.sig.pathways <- sig.results %>%
                        slice_max(order_by = Chi.square, n = 200) %>%
                        rownames()

ANOVA.sig.pathways <- sig.results2 %>%
                        slice_max(order_by = f.value, n = 200) %>%
                        rownames()

pathways.intersect <- intersect(KW.sig.pathways, ANOVA.sig.pathways)

#pathways.intersect <- pathways.intersect[-51]
pathways.intersect <- pathways.intersect[-c(3,4,8,41,47,57,62,73,78,80,85,86,90,99,101,120,125,139,141)]
pathways.intersect <- pathways.intersect[!grepl("REACTOME", pathways.intersect) ]


summary <- ES2 %>%
  group_by(seurat_clusters) %>%
  summarise(across(pathways.intersect, median))

normalize <- function(x) {
      (x- min(x)) /(max(x)-min(x))
}

for(i in 2:43) {
  summary[,i] <- normalize(summary[,i])
}
summary <- as.data.frame(summary)
#summary <- summary[,-5]
#summary <- summary[,-9]
#summary <- summary[,-80]

rownames(summary) <- summary$seurat_clusters
colnames(summary) <- stringr::str_remove_all(colnames(summary), "_UCell")

pdf("./output/Figure2/GSEA.heatmap.pdf", height = 8, width = 10)
pheatmap::pheatmap(t(summary[,2:43]))
dev.off()



################################
#Calculating PCA on Enrichment
###############################

x <- prcomp(ES2[,which(colnames(ES2) %in% pathways.intersect)], center = TRUE, scale. = TRUE)
eigs <- x$sdev^2
eigs <- eigs[1:10]

var_explained <- eigs/sum(eigs)

df <- as.data.frame(x$rotation[,1:10])
df$names <- rownames(df)
options(ggrepel.max.overlaps = Inf)
fviz_pca_var(x,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE ,    # Avoid text overlapping
             select.var = list(contrib = 60), 
             labelsize = 2)

ggsave("./output/Figure2/Overall.contribution.pdf", height = 20, width = 20)

fviz_contrib(x, choice = "var", axes = 1, top = 10)
ggsave("./output/Figure2/PC1_contribution.pdf", height = 7, width = 7)

fviz_contrib(x, choice = "var", axes = 2, top = 10)
ggsave("./output/Figure2/PC2_contribution.pdf", height = 7, width = 7)
    
df %>%
    ggplot(aes(x=df[,1],y=df[,2])) + 
    geom_point() + 
    geom_text(aes_string(label = "names"), 
            size=2, hjust = 0.5, nudge_y = -0.01) + 
    geom_hline(yintercept = 0, lty=2) + 
    geom_vline(xintercept = 0, lty=2) +
    labs(x=paste0("PC1",": ",round(var_explained[1]*100,1),"%"),
            y=paste0("PC2", ": ",round(var_explained[2]*100,1),"%")) +
    theme_classic()
ggsave("./output/Figure2/Figure2E.1_TFHonly.pdf", height = 7, width = 7)
    
df <- data.frame(x$x[,1:30], meta)

ggplot(df) +
  geom_hline(yintercept = 0, lty = 2) + 
  geom_vline(xintercept = 0, lty = 2) + 
  stat_density_2d(aes(x=PC1, y = PC2, fill = stat(level)), 
                  geom = "polygon", color = "black", lwd = 0.25, bins = 20) + 
  facet_grid(.~timepoint) + 
  labs(x=paste0("PC1",": ",round(var_explained[1]*100,1),"%"),
       y=paste0("PC2", ": ",round(var_explained[2]*100,1),"%")) +
  scale_fill_gradientn(colors = rev(brewer.pal(11, "RdYlBu"))) + 
  theme_classic() + 
  guides(fill = "none")
ggsave("./output/Figure2/Figure2E.2_TFHonly.pdf", height = 2.75, width = 25)

ggplot(ES2, aes(x=seurat_clusters, y = normalize(HALLMARK_TNFA_SIGNALING_VIA_NFKB_UCell))) + 
  geom_violin(aes(fill = seurat_clusters)) +
  geom_boxplot(width = 0.25) + 
  scale_fill_manual(values = mycolors[c(1:5,7:12)]) + 
  guides(fill = "none") + 
  theme_classic() + 
  theme(axis.title = element_blank())

ggsave("./output/Figure2/TNFsignaling.pdf", height = 2.75, width = 4)
```


```{r}
DimPlot(tfh, group.by = "cloneType") + 
  scale_color_manual(values = colorRampPalette(brewer.pal(11, "RdYlBu"))(4)[c(1,2,4)]) + 
  theme_void() + 
  theme(plot.title = element_blank())
ggsave("./output/Figure3/Figure2E.1_TFHonly.pdf", height = 3, width = 6)

df <- data.frame(tfh@reductions$umap@cell.embeddings, tfh[[]])

y <- df %>%
  group_by(orig.ident, CTaa) %>% 
  count()

y <- y %>%
  group_by(orig.ident) %>%
  mutate(sum.n = sum(n))

y$scaled <- y$n/y$sum.n

df$new.merge <- paste0(df$orig.ident, "_", df$CTaa)
y$new.merge <- paste0(y$orig.ident, "_", y$CTaa)


cloneTypes=c(Rare = 1e-4, Small = 0.001, 
                              Medium = 0.01, Large = 0.1, Hyperexpanded = 1)
cloneTypes <- c(None = 0, cloneTypes)
y$cloneType <- NA
for (x in seq_along(cloneTypes)) { names(cloneTypes)[x] <- 
        paste0(names(cloneTypes[x]), ' (', cloneTypes[x-1], 
        ' < X <= ', cloneTypes[x], ')') }

for (i in 2:length(cloneTypes)) { 
  y$cloneType <- ifelse(y$scaled > cloneTypes[i-1] & y$scaled <= cloneTypes[i], names(cloneTypes[i]), y$cloneType) 
}

colnames(y)[7] <- "Clone.Rep"
y <- y[,-c(1:2)]
df <- merge(df, y, by = "new.merge")

ggplot(df, aes(x= UMAP_1, y = UMAP_2)) + 
  geom_point(aes(color = Clone.Rep), size = 0.5) + 
  guides(color = "none") + 
  scale_color_manual(values = colorRampPalette(brewer.pal(11, "RdYlBu"))(7)[c(1,3,6)]) + 
  theme_void()
ggsave("./output/Figure2/Figure2E.1_TFHonly.pdf")

ggplot(df, aes(x= UMAP_1, y = UMAP_2)) + 
  geom_point(aes(color = sqrt(scaled)), size = 0.5, stroke = 0) + 
   guides(color = "none") + 
    scale_color_gradientn(colors = rev(brewer.pal(11, "RdYlBu"))) + 
  theme_void()
ggsave("./output/Figure2/Figure2E.1_TFHonly.png", height = 3.5, width = 4, dpi = 600, bg = "transparent")


tfh.28 <- subset(tfh, timepoint == "d28")

circles <- getCirclize(tfh.28, 
                       group.by = "ident")

compareClonotypes(tfh.28, 
                  cloneCall = "CTaa", 
                  split.by = "seurat_clusters", 
                  numbers = 10, 
                  samples = c(3,8)) 

#Just assigning the normal colors to each cluster
grid.cols <- colorRampPalette(brewer.pal(11, "Paired"))(length(unique(tfh$seurat_clusters)))
names(grid.cols) <- levels(tfh@active.ident)


#Graphing the chord diagram
pdf("./output/Figure2/tfh.d28.pdf", height = 4, width = 4)
circlize::chordDiagram(circles,
                       self.link = 1, 
                       grid.col = grid.cols, 
                       directional = 1,
                       direction.type =  "arrows",
                       link.arr.type = "big.arrow")
dev.off()

tfh.60 <- subset(tfh, timepoint == "d60")
circles <- getCirclize(tfh.60, 
                       group.by = "ident")

compareClonotypes(tfh.60, 
                  cloneCall = "CTaa", 
                  split.by = "seurat_clusters", 
                  numbers = 10, 
                  samples = c(3,8)) 

pdf("./output/Figure2/tfh.d60.pdf", height = 4, width = 4)
circlize::chordDiagram(circles,
                       self.link = 1, 
                       grid.col = grid.cols, 
                       directional = 1,
                       direction.type =  "arrows",
                       link.arr.type = "big.arrow")
dev.off()

tfh.110 <- subset(tfh, timepoint == "d110")
circles <- getCirclize(tfh.110, 
                       group.by = "ident")

compareClonotypes(tfh.110, 
                  cloneCall = "CTaa", 
                  split.by = "seurat_clusters", 
                  numbers = 10, 
                  samples = c(3,8)) 

pdf("./output/Figure2/tfh.d110.pdf", height = 4, width = 4)
circlize::chordDiagram(circles,
                       self.link = 1, 
                       grid.col = grid.cols, 
                       directional = 1,
                       direction.type =  "arrows",
                       link.arr.type = "big.arrow")
dev.off()


tfh.201 <- subset(tfh, timepoint == "d201")
compareClonotypes(tfh.201, 
                  cloneCall = "CTaa", 
                  split.by = "seurat_clusters", 
                  numbers = 10, 
                  samples = c(3,8)) 

circles <- getCirclize(tfh.201, 
                       group.by = "ident")


pdf("./output/Figure2/tfh.d201.pdf", height = 4, width = 4)
compareClonotypes(tfh.110, 
                  cloneCall = "CTaa", 
                  split.by = "seurat_clusters", 
                  numbers = 10, 
                  samples = c(3,8)) circlize::chordDiagram(circles,
                       self.link = 1, 
                       grid.col = grid.cols, 
                       directional = 1,
                       direction.type =  "arrows",
                       link.arr.type = "big.arrow")
dev.off()

```


#Figure 2F
```{r}
sampled <- df %>%
  group_by(timepoint) %>%
  sample_n(2000)

sampled$timepoint <- factor(sampled$timepoint, levels = c("d28", "d60", "d110", "d201"))
tfh$timepoint <- factor(tfh$timepoint, levels = c("d28", "d60", "d110", "d201"))
df <- data.frame(tfh[[]], tfh@reductions$umap@cell.embeddings)

xmax <- round(max(df$UMAP_1) + 3)
xmin <- round(min(df$UMAP_1) - 3)
ymax <- round(max(df$UMAP_2) + 3)
ymin <- round(min(df$UMAP_2) - 3)

ggplot(sampled, aes(x = UMAP_1, y = UMAP_2)) +
    geom_point(color = "grey") + 
     stat_density_2d(mapping = aes(fill = stat(level)),
                              geom = "polygon", 
                              bins = 20) + 
  geom_point(data = sampled[c(which(sampled$TCRA.S167.180 == "Specific"), which(sampled$TCRB.S167.180 == "Specific")), ], aes(x = UMAP_1, y = UMAP_2), fill = "white", shape = 21, size = 3) + 
          facet_grid(~timepoint) + 
           scale_fill_gradientn(colors = rev(brewer.pal(11, "RdYlBu"))) +
  ylim(ymin, ymax) + 
  xlim(xmin, xmax)+
  guides(fill = "none") + 
  theme_classic()
ggsave("./output/Figure2/Figure2F_TFHonly.pdf", height = 3, width = 10.5)

df2 <- df
df2$spike <- "No"
df2$spike[which(df2$TCRA.S167.180 == "Specific")] <- "Yes"
df2$spike[which(df2$TCRB.S167.180 == "Specific")] <- "Yes"

x <- df2 %>%
  group_by(timepoint, spike) %>%
  count()
```

#Figure 2G
```{r}
table <- table(sampled$seurat_clusters, sampled$timepoint)

for(x in seq_len(ncol(table))) {
  table[,x] <- table[,x]/sum(table[,x])
}
table <- data.frame(table)

ggplot(table, aes(x=Var1, y = Freq, fill = Var2)) + 
  geom_bar(stat = "identity", position = "fill") + 
  scale_fill_manual(values = colorRampPalette(brewer.pal(11, "RdYlBu"))(4)) + 
  xlab("Clusters") + 
  ylab("Relative Proportion") + 
  theme_classic()
ggsave("./output/Figure3/Figure2G_TFHonly.pdf", height = 3, width = 5)
```


################
#Figure 3 Graphs
################

```{r}
dir.create("./output/Figure3/")
```


# Loading and Filtering Data


################will need to add the old filter for TCR related Contigs

```{r}
library(dplyr)
library(Trex)
library(mumosa)

list <- readRDS("./data/processedData/Fullintegrated_seuratObjects_TFH.rds")
cells.to.remove <- unique(c(which(nchar(stringr::str_split(list$CTaa, "_", simplify = TRUE)[,1]) > 50), #TRA chain > 50 aa
                     which(nchar(stringr::str_split(list$CTaa, "_", simplify = TRUE)[,2]) > 50))) # TRB chain > 50 aa residues
if(length(cells.to.remove) > 0) {
    cells <- rownames(list[[]])
    cells <- cells[-cells.to.remove]
    list <- subset(list, cells = cells)
}

###################################################
#Generating Seurat Object for individual clonotypes
###################################################

conga <- CoNGAfy(list,
                 meta.carry = c("CTaa", "CTgene"))

meta <- list[[]]
clone.count <- meta %>%
  group_by(CTaa, TCRB.epitope, TCRA.epitope) %>%
  summarise(n = n()) %>%
  as.data.frame()

ID <- clone.count$CTaa
clone.count <- clone.count[,-1]
rownames(clone.count) <- ID

conga <- AddMetaData(conga, clone.count)
saveRDS(conga, "./data/processedData/conga_tfh_seuratObject.rds")

####################
#Adding Meta Data
####################

meta <- list[[]]

CTaa.unique <- unique(meta$CTaa)

for(i in seq_along(CTaa.unique)) {
  meta.tmp <- meta[meta$CTaa == CTaa.unique[i],]
  CTaa.tmp <- c(CTaa = unique(meta.tmp$CTaa), 
                donor = (paste(str_sort(unique(meta.tmp$donor)), collapse = ";")), 
                tissue = paste(str_sort(unique(meta.tmp$tissue)), collapse = ";"), 
                timepoint = paste(str_sort(unique(meta.tmp$timepoint)), collapse = ";"), 
                CD4.annot = paste(str_sort(unique(meta.tmp$CD4.annot)), collapse = ";"), 
                CD8.annot = paste(str_sort(unique(meta.tmp$CD8.annot)), collapse = ";"),
                DICE.annot = paste(str_sort(unique(meta.tmp$DICE.labels)), collapse = ";"), 
                CD8.score = paste(str_sort(unique(meta.tmp$CD8.pos)), collapse = ";"), 
                sequencing.run <- paste(str_sort(unique(meta.tmp$orig.ident)), collapse = ";"), 
                clusters <- paste(str_sort(unique(meta.tmp$seurat_clusters)), collapse = ";"))
  if(i == 1) {
    CTaa.Library <- CTaa.tmp
  } else {
    CTaa.Library <- rbind(CTaa.Library, CTaa.tmp)
  }
}
write.csv(CTaa.Library, "CTaa.tfh.library.csv")

CTaa.Library <- read.csv("CTaa.tfh.library.csv")
CTaa.Library <- CTaa.Library[,-1]
aa <- CTaa.Library$CTaa
CTaa.Library <- CTaa.Library[,-1]
rownames(CTaa.Library) <- aa
colnames(CTaa.Library)[8:9] <- c("SequencingRun", "Orig.Cluster")

conga <- AddMetaData(conga, CTaa.Library)

###################################
#Adding Test TCR sequence informtion
#####################################

confirming.seq <- list(TCRA = c("CATGLTGGGNKLTF", "CAGQLNRAAGNKLTF", "CIVPRGTGFQKLVF", "CAVRPNAGNNRKLIW", "CAVTGTYKYIF"), 
                       TCRB = c("CASSMGGGNQPQHF", "CASSPRGTEAFF", "CASSLVGGGPAEAFF", "CASSLGENDEQFF", "CASSQGNRANTEAFF")) 

conga$confirmation.seq <- NA
for (i in seq_len(5)) {
  x.pos <- intersect(grep(confirming.seq[[1]][i], conga$CTaa), grep(confirming.seq[[2]][i], conga$CTaa))
  conga$confirmation.seq[x.pos] <- paste0("TCR.", i)
}

conga$TRAV <- str_split(conga$CTgene, "[.]", simplify = TRUE)[,1]
conga$TRAV <- factor(conga$TRAV, levels = str_sort(unique(conga$TRAV), numeric = TRUE))
conga$TRBV <- str_split(str_split(conga$CTgene, "_", simplify = TRUE)[,2], "[.]", simplify = TRUE)[,1]
conga$TRBV <- factor(conga$TRBV, levels = str_sort(unique(conga$TRBV), numeric = TRUE))


#################################
#Performing Dimensional Reduction
#################################
dimen <- 15

conga <- conga %>% 
            NormalizeData(verbose = FALSE) %>% 
            FindVariableFeatures() 
conga <- Trex::quietTCRgenes(conga)
conga <- ScaleData(conga, verbose = FALSE) 
conga <- RunPCA(conga, features = VariableFeatures(conga))
         
dem.tmp <- c(1:dimen, 31:(31+dimen-1), 61:(61+dimen-1))

#Calculating mutual neighbor values
output <- rescaleByNeighbors(
                  list(conga[["TRA.KF"]]@cell.embeddings[,1:dimen], 
                      conga[["TRB.KF"]]@cell.embeddings[,1:dimen],
                      conga[["pca"]]@cell.embeddings[,1:dimen]), 
                  k = 100)

conga[["KF.corrected"]] <- CreateDimReducObject(
                                embeddings = as.matrix(output),
                                stdev = rep(0, ncol(output)),
                                key = "correctedTrex",
                                jackstraw = NULL,
                                misc = list())

#Calculating Phate embeding
phate.results <- phate(conga[["KF.corrected"]]@cell.embeddings[,1:length(dem.tmp)])


        
        
conga[["phateTrex"]] <- CreateDimReducObject(
                                        embeddings = phate.results$embedding,
                                        stdev = rep(0, 2),
                                        key = "phateTrex",
                                        jackstraw = NULL,
                                        misc = list())

#Calculating Neighbors
conga <- FindNeighbors(conga, 
                         reduction = "KF.corrected",
                         dims = 1:length(dem.tmp),
                         force.recalc = TRUE,
                         graph.name = c("T_nn", "T_snn"))

#Calculating Clusters
nn <- makeKNNGraph(conga@reductions$KF.corrected@cell.embeddings,
                                               k=20)
leiden.clusters <- leiden.community(nn, resolution = 0.7, n.iterations = 5)

conga$leiden.clusters <- leiden.clusters$membership
        

saveRDS(conga, "./data/processedData/conga_tfh_seuratObject.rds")
```


#Figure 3B
```{r}
conga <- readRDS("./data/processedData/conga_tfh_seuratObject.rds")
df <- data.frame(conga@reductions$phateTrex@cell.embeddings,
                  conga[[]])

centers = df %>% 
          group_by(leiden.clusters) %>% 
          summarize(across(c(paste0(dimRed[i], "_1"), paste0(dimRed[i], "_2")),mean)) %>%
          as.data.frame()
        max.cluster.val <- max(as.numeric(na.omit(df$leiden.clusters)))
        df$leiden.clusters <- factor(df$leiden.clusters, levels = c(0:max.cluster.val))

mycolors <- colorRampPalette(brewer.pal(12,"Paired"))(length(unique(df$leiden.clusters)))

ggplot() +
    geom_point(data =df, aes(x=phateTrex_1, y = phateTrex_2, size = n, fill = leiden.clusters), 
               shape =21, stroke = 0.1) + 
    scale_fill_manual(values = mycolors) + 
    theme_classic() +
    geom_text(data = centers, mapping =  aes(x=centers[,2], y = centers[,3], label = leiden.clusters), 
              color = "black", size = 5) +
    labs(fill = "cluster") + 
    theme(axis.title = element_blank())

ggsave("./output/Figure3/Figure3B.png", height = 4, width = 5.75, dpi = 600)
```

#Figure 3C
```{r}
x.features <- c("BCL6", "CXCL13", "CXCR5", "ICOS", "MAF", "PDCD1", "EMP3", "KLF2")
x <- FeaturePlot(conga, 
              features = x.features,
              reduction = "phateTrex",
            cols = rev(brewer.pal(11, "RdYlBu")), 
            coord.fixed = FALSE, 
            combine = FALSE) 
for(i in seq_along(x)) {
  x[[i]] <- x[[i]] + 
    xlim(-0.03,0.05) +
    ylim(-0.01,0.015) +
    guides(color = "none") + 
    theme_void()
}
plot_a_list(x,3,2)
ggsave("./output/Figure3/Figure3C.png", height = 10.67, width = 6.67, dpi = 600)
```

#Figure 3D
```{r}
epitope <- stringr::str_sort(na.omit(unique(df$TCRA.epitope)), numeric = TRUE)
        
ep.colors <- rev(brewer.pal(5, "RdYlBu"))[c(1,3,5)]
names(ep.colors) <- epitope

df$TCRA.epitope <- factor(df$TCRA.epitope, levels = epitope)
  
# Plotting the TCRA-specific sequences
df.sub <- subset(df, !is.na(TCRA.epitope))
ggplot() +
  geom_point(data =df, aes(x=phateTrex_1, y = phateTrex_2, size = n), 
             fill = "grey", shape =21, stroke = 0.1) + 
  geom_point(data = df.sub, aes(x=phateTrex_1, y = phateTrex_2, size = n, fill = TCRA.epitope), 
             shape =21, stroke = 0.1) + 
  scale_fill_manual(values = ep.colors) + 
  guides(size = "none", fill = "none") + 
  theme_void()
  ggsave("./output/Figure3/Figure3D_TRA.png", height = 4, width = 5, dpi = 600)

# Plotting the TCRB-specific sequences
df$TCRB.epitope <- factor(df$TCRB.epitope, levels = epitope)

df.sub <- subset(df, !is.na(TCRB.epitope))
ggplot() +
    geom_point(data =df, aes(x=phateTrex_1, y = phateTrex_2, size = n), 
               fill = "grey", shape =21, stroke = 0.1) + 
    geom_point(data = df.sub, aes(x=phateTrex_1, y = phateTrex_2, size = n, fill = TCRB.epitope), 
               shape =21, stroke = 0.1) + 
    scale_fill_manual(values = ep.colors) + 
    guides(size = "none", fill = "none") + 
    theme_void()

ggsave("./output/Figure3/Figure3D_TRB.png", height = 4, width = 5, dpi = 600)
```

#Figure 3E

```{r}
library(seqinr)

clusters <- c(0,1,3)
conga.meta <- conga[[]]
for(i in seq_along(clusters)) {
  tmp.meta <- conga.meta[conga.meta$leiden.clusters == clusters[i],]
  tmp.meta <- tmp.meta[!is.na(tmp.meta$TCRB.epitope) |!is.na(tmp.meta$TCRA.epitope),] 
  TRB <- str_split(tmp.meta$CTaa, "_", simplify = TRUE)[,2]
  TRA <- str_split(tmp.meta$CTaa, "_", simplify = TRUE)[,1]
  align <- msa(TRA, 
               type = "protein", 
               maxiters = 30)
  align2 <- msaConvert(align, type="seqinr::alignment")
  
  ggseqlogo(align2$seq, 
            seq_type='aa', 
            method='p') 
  ggsave(paste0("./output/Figure3/Figure3E.Cluster.", clusters[i], ".TRA.pdf"), height =3, width = 4.5)
  
    align <- msa(TRB, 
               type = "protein", 
               maxiters = 30)
  align2 <- msaConvert(align, type="seqinr::alignment")
  
  ggseqlogo(align2$seq, 
            seq_type='aa', 
            method='p') 
  ggsave(paste0("./output/Figure3/Figure3E.Cluster.", clusters[i], ".TRB.pdf"), height =3, width = 4.5)
}
```


#Figure 3F
```{r}
df.sub <- subset(df, !is.na(confirmation.seq))

ggplot() +
    geom_point(data =df, aes(x=phateTrex_1, y = phateTrex_2, size = n), 
               fill = "grey", shape =21, stroke = 0.1) + 
    geom_point(data = df.sub, aes(x=phateTrex_1, y = phateTrex_2, fill = confirmation.seq), 
               size = 4, shape =21, stroke = 0.2) + 
    scale_fill_manual(values = brewer.pal(5, "RdYlBu")) +
    theme_void() + 
    guides(fill = "none", size = "none")

ggsave("./output/Figure3/Figure3F.png", height = 4, width = 5, dpi = 600)
```

############
#Comparing the cluster output

```{r}
library(ggthemes)
conga <- RunUMAP(conga,
                 reduction = "pca",
                 dims = 1:25,
                 n.epochs = 500, 
                 reduction.key = "rnaUMAP_")

nn <- makeKNNGraph(conga@reductions$TRA.KF@cell.embeddings, k=20)
leiden.clusters <- leiden.community(nn, resolution = 0.7, n.iterations = 5)
conga$TRA.clusters <- leiden.clusters$membership

nn <- makeKNNGraph(conga@reductions$TRB.KF@cell.embeddings, k=20)
leiden.clusters <- leiden.community(nn, resolution = 0.7, n.iterations = 5)
conga$TRB.clusters <- leiden.clusters$membership

nn <- makeKNNGraph(conga@reductions$pca@cell.embeddings, k=20)
leiden.clusters <- leiden.community(nn, resolution = 0.7, n.iterations = 5)
conga$RNA.clusters <- leiden.clusters$membership



hclust.cluster <- read.csv("./processedData/hclust.TCRdist.leiden.csv")[,2:3]
knn.cluster <- read.csv("./processedData/knn.TCRdist.leiden.csv")[,2:3]
alpha.mc <- read.csv("./processedData/alphaMC.TCRdist.leiden.csv")[,2:3]

clusters <- merge(hclust.cluster, knn.cluster, by = 2, all.x = TRUE)
clusters <- merge(clusters, alpha.mc, by.x = 1, by.y = 2, all.x = TRUE)
clones <- clusters$clonotype
clusters <- clusters[,-1]
rownames(clusters) <- clones

conga <- AddMetaData(conga, clusters)


df <- data.frame(conga@reductions$umap@cell.embeddings,
                  conga[[]])

df.sub <- subset(df, !is.na(confirmation.seq))

ggplot() +
    geom_point(data =df, aes(x=rnaUMAP_1, y = rnaUMAP_2, size = n), 
               fill = "grey", shape =21, stroke = 0.1) + 
    geom_point(data = df.sub, aes(x=rnaUMAP_1, y = rnaUMAP_2, fill = confirmation.seq), 
               size = 4, shape =21, stroke = 0.2) + 
    scale_fill_manual(values = rev(brewer.pal(5, "RdYlBu"))) +
    theme_void() + 
    guides(fill = "none", size = "none")
ggsave("./output/TCRsequences.cluster.png", height = 3, width = 4.25, dpi = 600)


ggplot() +
    geom_point(data =df, aes(x=rnaUMAP_1, y = rnaUMAP_2, size = n, fill = RNA.clusters), 
               shape =21, stroke = 0.1) + 
    theme_void() + 
    scale_fill_tableau()
ggsave("./output/RNA.cluster.png", height = 3, width = 4.25, dpi = 600)

mycolors <- colorRampPalette(brewer.pal(12,"Paired"))(length(unique(df$leiden.clusters)))
ggplot() +
    geom_point(data =df, aes(x=rnaUMAP_1, y = rnaUMAP_2, size = n, fill = leiden.clusters), 
               shape =21, stroke = 0.1) + 
    theme_void() + 
    scale_fill_manual(values = mycolors) + 
  labs(fill='Trex.Cluster')
ggsave("./output/Trex.cluster.png", height = 3, width = 4.25, dpi = 600)


ggplot() +
    geom_point(data =df, aes(x=rnaUMAP_1, y = rnaUMAP_2, size = n, fill = as.factor(hclust.TCRdist.leiden)), 
               shape =21, stroke = 0.1) + 
    theme_void() + 
    scale_fill_viridis(option = "H", discrete = TRUE) + 
  labs(fill='TCRdist.Hclust')
ggsave("./output/Hclust.cluster.png", height = 3, width = 4.35, dpi = 600)


ggplot() +
    geom_point(data =df, aes(x=rnaUMAP_1, y = rnaUMAP_2, size = n, fill = as.factor(knn.TCRdist.leiden)), 
               shape =21, stroke = 0.1) + 
    theme_void() + 
    scale_fill_viridis(option = "B", discrete = TRUE, na.value = "grey") + 
    guides(fill = "none")
ggsave("./output/Knn.cluster.png", height = 3, width = 4.25, dpi = 600)


ggplot() +
    geom_point(data =df, aes(x=rnaUMAP_1, y = rnaUMAP_2, size = n, fill = as.factor(alphaMC.TCRdist.leiden)), 
               shape =21, stroke = 0.1) + 
    theme_void() + 
    scale_fill_viridis(option = "B", discrete = TRUE, na.value = "grey") + 
    guides(fill = "none")
ggsave("./output/AlphaMC.cluster.png", height = 3, width = 4.25, dpi = 600)

```


```{r}
conga <- annotateDB(conga, 
                    chains = "TRB")
conga <- annotateDB(conga, 
                    chains = "TRA")



conga$TRA_Epitope.species <- stringr::str_remove(conga$TRA_Epitope.species, "NA;")
conga$TRA_Epitope.species <- stringr::str_remove(conga$TRA_Epitope.species, ";NA")
conga$TRA_Epitope.species <- stringr::str_remove(conga$TRA_Epitope.species, "HGNC_Genes;")
conga$TRA_Epitope.species <- stringr::str_remove(conga$TRA_Epitope.species, ";HGNC_Genes")
conga$TRA_Epitope.species <- stringr::str_remove(conga$TRA_Epitope.species, "HGNC_Genes")

conga$TRA_Epitope.species <- stringr::str_remove(conga$TRA_Epitope.species, "Malignant neoplasm of brain;")
conga$TRA_Epitope.species <- stringr::str_remove(conga$TRA_Epitope.species, ";Malignant neoplasm of brain")
conga$TRA_Epitope.species <- stringr::str_remove(conga$TRA_Epitope.species, "Malignant neoplasm of brain")
conga$TRA_Epitope.species[which(conga$TRA_Epitope.species == "")] <- NA


conga$CMV <- NA
conga$CMV[grepl("CMV", conga$TRA_Epitope.species)] <- "CMV"
conga$CMV[grepl("CMV", conga$TRB_Epitope.species)] <- "CMV"

conga$Influenza <- NA
conga$Influenza[grepl("InfluenzaA|Influenza", conga$TRA_Epitope.species)] <- "Influenza"
conga$Influenza[grepl("InfluenzaA|Influenza", conga$TRB_Epitope.species)] <- "Influenza"

conga$EBV <- NA
conga$EBV[grepl("EBV", conga$TRA_Epitope.species)] <- "EBV"
conga$EBV[grepl("EBV", conga$TRB_Epitope.species)] <- "EBV"

conga$SARSCoV2 <- NA
conga$SARSCoV2[grepl("SARS-CoV-2", conga$TRA_Epitope.species)] <- "SARSCoV2"

conga$M.tuberculosis <- NA
conga$M.tuberculosis[grepl("M. tuberculosis|M.Tuberculosis|M.tuberculosis", conga$TRA_Epitope.species)] <- "M.tuberculosis"
conga$M.tuberculosis[grepl("M. tuberculosis|M.Tuberculosis|M.tuberculosis", conga$TRB_Epitope.species)] <- "M.tuberculosis"

conga$Autoimmune <- NA
conga$Autoimmune[grepl("Psoriatic arthritis|Multiple sclerosis|Celiac disease|Inflammatory bowel disease|Systemic lupus erythematosus|Thyrotoxicosis with diffuse goitre|Churg Strauss syndrome|Diabetes Type 1|Multiple sclerosis (MS)", conga$TRA_Epitope.species)] <- "Autoimmune"

conga$Autoimmune[grepl("Psoriatic arthritis|Multiple sclerosis|Celiac disease|Inflammatory bowel disease|Systemic lupus erythematosus|Thyrotoxicosis with diffuse goitre|Churg Strauss syndrome|Diabetes Type 1|Multiple sclerosis (MS)", conga$TRB_Epitope.species)] <- "Autoimmune"


df <- data.frame(conga@reductions$phateTrex@cell.embeddings,
                  conga[[]])

df <- df[,c("phateTrex_1", "phateTrex_2", "CMV", "SARSCoV2", "M.tuberculosis", "EBV", "Influenza", "Autoimmune")]
df <- reshape2::melt(df, id.vars = c("phateTrex_1", "phateTrex_2"))

ggplot() +
 geom_point(data =df, aes(x=phateTrex_1, y = phateTrex_2), 
             fill = "grey", shape =21, stroke = 0.1) + 
  stat_density_2d(data = subset(df, variable == "Autoimmune" & !is.na(df$value)), aes(x =phateTrex_1, y = phateTrex_2, fill = stat(level)), 
                  geom = "polygon", bins = 25) + 
  stat_density_2d(data = subset(df, variable == "CMV" & !is.na(df$value)), aes(x =phateTrex_1, y = phateTrex_2, fill = stat(level)), 
                  geom = "polygon", bins = 25) + 
  stat_density_2d(data = subset(df, variable == "SARSCoV2" & !is.na(df$value)), aes(x =phateTrex_1, y = phateTrex_2, fill = stat(level)), 
                  geom = "polygon", bins = 25) + 
  stat_density_2d(data = subset(df, variable == "EBV" & !is.na(df$value)), aes(x =phateTrex_1, y = phateTrex_2, fill = stat(level)), 
                  geom = "polygon", bins = 25) + 
  stat_density_2d(data = subset(df, variable == "Influenza" & !is.na(df$value)), aes(x =phateTrex_1, y = phateTrex_2, fill = stat(level)), 
                  geom = "polygon", bins = 25) + 
  stat_density_2d(data = subset(df, variable == "M.tuberculosis" & !is.na(df$value)), aes(x =phateTrex_1, y = phateTrex_2, fill = stat(level)), 
                  geom = "polygon", bins = 25) + 
  scale_fill_gradientn(colors = rev(brewer.pal(11, "RdYlBu"))) + 
  facet_wrap(~variable) +
  guides(fill = "none") + 
  xlim(-0.02,0.05) + 
  ylim(-0.015,0.015) + 
  theme_void() 
ggsave("Epitope.sample.pdf", height = 5, width = 7.5)

```

